---
title: "Population Simulation"
output: 
  html_document:
      theme: united
      css: style.css
params:
  n: 100000
  dim: 10 
  rho: 0.1
  thresh: 0.2
  eps_sd: 1
    
  r: 1 
  initPopSize: 500 
  carryingCapacity: 500 
  timeSteps: 10000 
  
  setseed: FALSE
  seed: 10
  
  filename: "testtext"
  fileext: ".txt"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, cache = F, fig.align = 'center')
if(params$setseed){
  set.seed(params$seed)
}
```

# Initial Values
These are the parameters to construct our copula with tail dependence for which we obtain our environmental noise.

* <b>n:</b> `r params$n`<br />
* <b>dim:</b> `r params$dim`<br />
* <b>rho:</b> `r params$rho`<br />
* <b>thresh:</b> `r params$thresh`<br />
* <b>eps_sd:</b> `r params$eps_sd`<br />

These are the values that we have when we simulate our population throughout time.

* <b>r:</b> `r params$r`<br />
* <b>numSimulations:</b> `r params$numSimulations`<br />
* <b>initPopSize:</b> `r params$initPopSize`<br />
* <b>carryingCapacity:</b> `r params$carryingCapacity`<br />
* <b>timeSteps:</b> `r params$timeSteps`<br />

This is if we set a seed for our program.

* <b>setseed:</b> `r params$setseed`
* <b>seed:</b> `r params$seed`

This is where we obtain additional information we want to include in our document.

* <b>filename:</b> `r params$filename` <br />
* <b>fileext:</b> `r params$fileext` <br />


```{r, message = F,echo=F}
library(copula)
library(mvtnorm)
library(ggplot2)
library(tseries)
library(forecast)
```

```{r, functions,echo=F}
#dr. reuman's function
newCopulaWithTailDependece<-function(U,res,n,dim,rho,thresh_lo,thresh_hi)
{
  res<-(thresh_hi-thresh_lo)*res+thresh_lo
  res[U>thresh_hi,]<-rep(U[U>thresh_hi],times=dim) #this line takes the rows of res for which the corresponding
  #entry of U is bigger than thresh_hi and replaces both entries of those rows by the corresponding value of U
  res[U<thresh_lo,]<-rep(U[U<thresh_lo],times=dim) #this line takes the rows of res for which the corresponding
  #entry of U is smaller than thresh_lo and replaces both entries of those rows by the corresponding value of U
  return(res)
}

newCopulaWithTail<- function(res,n,dim,rho,thresh_lo,thresh_hi){
  n1 <- round(thresh_lo*n)
  n2 <- round(thresh_hi*n)
  
}

# function isn't really needed, can avoid using it
findNormalCopulaParam <- function(dim,varSumRetain){
    fOptim <-function(rho){
     #varSum is the sum of all entries in the covariance matrix 
     # and the below formula gets that sum in a closed formula way
     varSum <- dim*(dim-1)*rho + dim
     return(abs(varSum-varSumRetain))
    }
    # $minimum is the x value for which the minimum is achieved (note that optimize returns a list) 
    return(optimize(fOptim,c(0,1),lower=0,upper=1,maximum=FALSE)$minimum)
}
```

# Construction of a Tail Dependent Random Variable
Here we construct our copula with tail dependence with the newCopulaWithTailDependence function.
```{r}
U<-runif(params$n)
ncop<-normalCopula(params$rho,dim=params$dim) 
res<-rCopula(params$n,ncop)
upper<-newCopulaWithTailDependece(U,res,params$n,params$dim,params$rho,0,1-params$thresh)
lower<-newCopulaWithTailDependece(U,res,params$n,params$dim,params$rho,params$thresh,1)
both<-newCopulaWithTailDependece(U,res,params$n,params$dim,params$rho,params$thresh/2,1-(params$thresh/2))
plot(upper)
plot(lower)
plot(both)
```

The parameters that we pass into the function indicate that:

* we sample from `r params$n` points from a copula 
* our copula has dimension `r params$dim` 
* we use the value of `r params$rho` for rho where rho is the normal copula parameter for the region of the copula that is not tail dependent.
* we have the values of `r params$thresh_lo` for thresh_lo and `r params$thresh_hi` for thresh_hi.


# Transformation of our Random Variable and Construction of New Random Variables
We now transforms all of our marginals of our `r params$dim`-dimensional copula with into normal marginals.
```{r}
upper<-qnorm(upper,sd=params$eps_sd)
lower<-qnorm(lower,sd=params$eps_sd)
both<-qnorm(both,sd=params$eps_sd)
cov(upper)
cov(lower)
cov(both)
```

`r con=file(paste0(getwd(),"/testhtml/",params$filename,params$fileext));lines=readLines(con,n=-1);close(con)`
<!-- `r lines` -->
We now find the rho parameter to construct a `r params$dim`-variate normal copula that has the same value for the  variance of the sum of the marginals as our tail dependent random variable
```{r}
varSumRetain <- sum((1/params$eps_sd^2)*cov(upper))
rho <- findNormalCopulaParam(params$dim, varSumRetain)
newNormalCop <- normalCopula(param = rho,dim =params$dim)
newNormCop <- rCopula(params$n, newNormalCop)
```
We new transform this normal copula into normal marginals.
```{r}
multiNorm <- qnorm(newNormCop,sd=params$eps_sd)
```
We now create our comonotonic random variable with dimension `r params$dim`
```{r}
temp <- rnorm(params$n, sd=params$eps_sd)
com <- matrix(data=replicate(n=params$dim,temp),ncol=params$dim)
```

```{r,echo=F}
sum_norm <- rowSums(multiNorm)
sum_upper <- rowSums(upper)
sum_lower <- rowSums(lower)
sum_both <- rowSums(both)
sum_com <-rowSums(com)
```

`r con=file(paste0(getwd(),"/testhtml/",params$filename,params$fileext));lines=readLines(con,n=-1);close(con)`
`r lines`
This is the empirical cumulative density function of three random
variables along with the cases of comonotonic and not comonotonic. These three random variables are the sum of the
`r params$dim` marginals of each the three cases. The variance of the sum of these
random variables respectively are:

* <b>comonotonic:</b> `r sum(cov(com))` `r var(sum_com)` 
* <b>comonotonic</b> in the upper tail: `r sum(cov(upper))` `r var(sum_upper)` 
* <b>comonotonic</b> in the lower tail: `r sum(cov(lower))` `r var(sum_lower)`
* <b>comonotonic</b> in both tails: `r sum(cov(both))` `r var(sum_both)`
* <b>not comonotonic:</b> `r sum(cov(multiNorm))` `r var(sum_norm)`

The following is the variance of the first marginal of each of the cases:

* <b>comonotonic:</b> `r var(com[,1])`
* <b>comonotonic</b> in the upper tail: `r var(upper[,1])`
* <b>comonotonic</b> in the lower tail: `r var(lower[,1])`
* <b>comonotonic</b> in both tails: `r var(both[,1])`
* <b>not comonotonic:</b> `r var(multiNorm[,1])`

This is the ecdfplot of the sum of the marginals of each of the three random variables.
```{r, ecdfplot,echo=F}
valuesV <- c(sum_com, sum_upper,sum_lower,sum_both,sum_norm)
typeV <- rep("sum_com",params$n)
typeV <- append(typeV, rep("sum_upper",params$n))
typeV <- append(typeV, rep("sum_lower",params$n))
typeV <- append(typeV, rep("sum_both",params$n))
typeV <- append(typeV, rep("sum_norm",params$n))


dataFrame <- data.frame(values = valuesV, type = typeV)

ecdfPlot <- ggplot(dataFrame, aes( x= unlist(dataFrame[1]) ) )
ecdfPlot <- ecdfPlot + stat_ecdf(mapping=aes(color=type))
ecdfPlot <- ecdfPlot + labs(title = "Empirical Cumulative \n Density Function")
ecdfPlot <- ecdfPlot +  xlab("Random variable value")
ecdfPlot <- ecdfPlot +  ylab("Cumulative Density")
ecdfPlot <- ecdfPlot + theme(text = element_text(size=7),
                             legend.box='horizontal',
                             legend.text = element_text(size = 8),
                             panel.background = element_rect(fill = "white")) + scale_colour_brewer(palette = "Set1")
plot(ecdfPlot)
```

`r con=file(paste0(getwd(),"/testhtml/",params$filename,params$fileext));lines=readLines(con,n=-1);close(con)`
`r lines`
We have three random variables where each of them have `r params$dim` marginals. Each of these random variables
represents the environmental noise on a population each with `r params$dim` metapopulations. We use a stochastic 
ricker model to model how the population fluctuates over time

```{r}
ricker <-function(curpop,r,k,eps){
  return(curpop*exp(r*(1-(curpop/k))+eps))
}
```


```{r, epsilon environment var, eval = T,echo=F}
eps_upper <- upper[sample(nrow(upper),size=params$timeSteps,replace=T),] 
eps_lower <- lower[sample(nrow(lower),size=params$timeSteps,replace=T),] 
eps_both <- both[sample(nrow(both),size=params$timeSteps,replace=T),] 
eps_com <- com[sample(nrow(com),size=params$timeSteps,replace=T),] 
eps_norm <- multiNorm[sample(nrow(multiNorm),size=params$timeSteps,replace=T),] 

popSizeCom <- matrix(data=rep(params$initPopSize,times=params$dim),nrow=1,byrow = T)
popSizeUpper <- matrix(data=rep(params$initPopSize,times=params$dim),nrow=1,byrow = T)
popSizeLower <- matrix(data=rep(params$initPopSize,times=params$dim),nrow=1,byrow = T)
popSizeBoth <- matrix(data=rep(params$initPopSize,times=params$dim),nrow=1,byrow = T)
popSizeNorm <- matrix(data=rep(params$initPopSize,times=params$dim),nrow=1,byrow = T)

for(i in 1:(params$timeSteps)){
    popSizeCom <- rbind(popSizeCom,ricker(popSizeCom[i,],params$r,params$carryingCapacity,eps_com[i,]))
    popSizeUpper <- rbind(popSizeUpper,ricker(popSizeUpper[i,],params$r,params$carryingCapacity,eps_upper[i,]))
    popSizeLower <- rbind(popSizeLower,ricker(popSizeLower[i,],params$r,params$carryingCapacity,eps_lower[i,]))
    popSizeBoth <- rbind(popSizeBoth,ricker(popSizeBoth[i,],params$r,params$carryingCapacity,eps_both[i,]))
    popSizeNorm <- rbind(popSizeNorm,ricker(popSizeNorm[i,],params$r,params$carryingCapacity,eps_norm[i,]))
}

avgPopSizeCom <- rowSums(popSizeCom)/params$dim 
avgPopSizeUpper <- rowSums(popSizeUpper)/params$dim 
avgPopSizeLower <- rowSums(popSizeLower)/params$dim
avgPopSizeBoth <- rowSums(popSizeBoth)/params$dim
avgPopSizeNorm <- rowSums(popSizeNorm)/params$dim

plot.ts(avgPopSizeCom)
plot.ts(avgPopSizeUpper)
plot.ts(avgPopSizeLower)
plot.ts(avgPopSizeBoth)
plot.ts(avgPopSizeNorm)

plot.ts(log10(avgPopSizeCom))
plot.ts(log10(avgPopSizeUpper))
plot.ts(log10(avgPopSizeLower))
plot.ts(log10(avgPopSizeBoth))
plot.ts(log10(avgPopSizeNorm))

if(params$dim >= 10){
  plot.ts(popSizeCom[,1:10])
  plot.ts(popSizeUpper[,1:10])
  plot.ts(popSizeLower[,1:10])
  plot.ts(popSizeBoth[,1:10])
  plot.ts(popSizeNorm[,1:10])
}
```

`r con=file(paste0(getwd(),"/testhtml/",params$filename,params$fileext));lines=readLines(con,n=-1);close(con)`
`r lines`

```{r}
upB <- max(quantile(avgPopSizeCom,0.98),quantile(avgPopSizeUpper,0.98),quantile(avgPopSizeLower,0.98),quantile(avgPopSizeBoth,0.98),quantile(avgPopSizeNorm,0.98))

c <- data.frame(population =avgPopSizeCom[avgPopSizeCom<upB] )
u <- data.frame(population =avgPopSizeUpper[avgPopSizeUpper<upB] )
l <- data.frame(population =avgPopSizeLower[avgPopSizeLower<upB] )
b <- data.frame(population =avgPopSizeBoth[avgPopSizeBoth<upB] )
n <- data.frame(population =avgPopSizeNorm[avgPopSizeNorm<upB] )

c$type <- "comonotonic"
u$type <- "comonotonic in the upper tail"
l$type <- "comonotonic in the lower tail"
b$type <- "comonotonic in both tails"
n$type <- "not comonotonic"

popfreq <- rbind(c,u,l,b,n)

ggplot(popfreq,aes(population,fill=type)) + geom_histogram(bins=200, alpha =0.6, position ='identity')

hist(avgPopSizeCom[avgPopSizeCom<upB],breaks=200)
print(paste0("mean: ", mean(avgPopSizeCom)," median: ", median(avgPopSizeCom)))

hist(avgPopSizeUpper[avgPopSizeUpper<upB],breaks=200)
print(paste0("mean: ", mean(avgPopSizeUpper)," median: ", median(avgPopSizeUpper)))

hist(avgPopSizeLower[avgPopSizeLower<upB],breaks=200)
print(paste0("mean: ", mean(avgPopSizeLower)," median: ", median(avgPopSizeLower)))

hist(avgPopSizeBoth[avgPopSizeBoth<upB],breaks=200)
print(paste0("mean: ", mean(avgPopSizeBoth)," median: ", median(avgPopSizeBoth)))

hist(avgPopSizeNorm[avgPopSizeNorm<upB],breaks=200)
print(paste0("mean: ", mean(avgPopSizeNorm)," median: ", median(avgPopSizeNorm)))

```

