---
title: "population simulation"
output: html_document
params:
  n: 10000
  dim: 10 
  rho: 0.1
  thresh_hi: 0.7
  thresh_lo: 0.0
  
  r: 1 
  numSimulations: 1
  initPopSize: 500 
  carryingCapacity: 500 
  timeSteps: 20000
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r}
library(copula)
library(mvtnorm)
library(ggplot2)
```

```{r, functions}
#dr. reuman's function
newCopulaWithTailDependece<-function(n,dim,rho,thresh_lo,thresh_hi)
{
  U<-runif(n)
  ncop<-normalCopula(rho,dim=dim) #this is the copula of a multivariate normal with covariance matrix having 1s on
  #the diagonal and all rhos in all off-diagonal positions
  res<-rCopula(n,ncop)
  res<-(thresh_hi-thresh_lo)*res+thresh_lo
  res[U>thresh_hi,]<-rep(U[U>thresh_hi],times=dim) #this line takes the rows of res for which the corresponding
  #entry of U is bigger than thresh_hi and replaces both entries of those rows by the corresponding value of U
  res[U<thresh_lo,]<-rep(U[U<thresh_lo],times=dim) #this line takes the rows of res for which the corresponding
  #entry of U is smaller than thresh_lo and replaces both entries of those rows by the corresponding value of U
  return(res)
}


findNormalCopulaParam <- function(dim,varSumRetain){
    fOptim <-function(rho){
     #varSum is the sum of all entries in the covariance matrix 
     # and the below formula gets that sum in a closed formula way
     varSum <- dim*(dim-1)*rho + dim
     return(abs(varSum-varSumRetain))
    }
    # $minimum is the x value for which the minimum is achieved (note that optimize returns a list) 
    return(optimize(fOptim,c(0,1),lower=0,upper=1,maximum=FALSE)$minimum)
}
```
## initial parameter values
params\$n: `r params$n`<br />
params\$dim: `r params$dim`<br />
params\$rho: `r params$rho`<br />
params\$thresh_hi: `r params$thresh_hi`<br />
params\$thresh_lo: `r params$thresh_lo`<br />

params\$r: `r params$r`<br />
params\$numSimulations: `r params$numSimulations`<br />
params\$initPopSize: `r params$initPopSize`<br />
params\$carryingCapacity: `r params$carryingCapacity`<br />
params\$timeSteps: `r params$timeSteps`<br />

```{r}
res<-newCopulaWithTailDependece(params$n,params$dim,params$rho,params$thresh_lo,params$thresh_hi)
plot(res[,1],res[,2])
res<-qnorm(res)
hist(res[,1],50)
hist(res[,2],50)
plot(res[,1],res[,2])
```

```{r, random variables}

covMatrix <- matrix(data=rep(0,times=params$dim*params$dim),nrow=params$dim,byrow=TRUE)
for(i in 1:params$dim){
  for(j in 1:(params$dim)){
   covMatrix[i,j] <- cov(res[,i],res[,j])
  }
}
varSumRetain <- sum(covMatrix)


rho <- findNormalCopulaParam(params$dim, varSumRetain)
newNormalCop <- normalCopula(param = rho,dim =params$dim)
newNormCop <- rCopula(params$n, newNormalCop)
newMultiNormMarg <- qnorm(newNormCop)
com_mat <- matrix(data=replicate(n=params$dim,res[,1]),ncol=params$dim)


new_sum_res <- rowSums(newMultiNormMarg)
sum_res <- rowSums(res)
sum_com <- params$dim*res[,1] # or rowSums(com_mat) is equivalent

hist(sum_com,50)
var(sum_com)
hist(sum_res,50)
var(sum_res)
hist(new_sum_res,50)
var(new_sum_res)
```

```{r, ecdfplot}
valuesV <- append(append(sum_com, sum_res),new_sum_res)
typeV <- rep("sum_com",params$n)
typeV <- append(typeV, rep("sum_res",params$n))
typeV <- append(typeV, rep("new_sum_res",params$n))


dataFrame <- data.frame(values = valuesV, type = typeV)

ecdfPlot <- ggplot(dataFrame, aes( x= unlist(dataFrame[1]) ) )
ecdfPlot <- ecdfPlot + stat_ecdf(mapping=aes(color=type))
ecdfPlot <- ecdfPlot + labs(title = "Empirical Cumulative \n Density Function")
ecdfPlot <- ecdfPlot +  xlab("Random variable value")
ecdfPlot <- ecdfPlot +  ylab("Cumulative Density")
ecdfPlot <- ecdfPlot + theme(text = element_text(size=7),
                             legend.box='horizontal',
                             legend.text = element_text(size = 8),
                             panel.background = element_rect(fill = "white")) + scale_colour_brewer(palette = "Set1")
plot(ecdfPlot)
```

```{r, epsilon environment var, warning = FALSE}
ricker <-function(curpop,r,k,eps){
  return(curpop*exp(r*(1-(curpop/k))+eps))
}

extinct<-function(pop,tLow,tHigh){
  if(any(pop<tLow|pop>tHigh)){return(TRUE)}
  return(FALSE)
}


tLow <- params$initPopSize/2 
tHigh <- 2*params$carryingCapacity

numExtinctCom <- 0 
numExtinctRes <- 0 
numExtinctNewRes <- 0 

for(j in 1:params$numSimulations){
  eps_res <- res[sample(nrow(res),size=params$timeSteps,replace=T),] 
  eps_com <- com_mat[sample(nrow(res),size=params$timeSteps,replace=T),] 
  eps_new_res <- newMultiNormMarg[sample(nrow(newMultiNormMarg),size=params$timeSteps,replace=T),] 
  
  popSizeCom <- matrix(data=rep(params$initPopSize,times=params$dim),nrow=1,byrow = T)
  popSizeRes <- matrix(data=rep(params$initPopSize,times=params$dim),nrow=1,byrow = T)
  popSizeNewRes <-  matrix(data=rep(params$initPopSize,times=params$dim),nrow=1,byrow = T)
  
  for(i in 1:(params$timeSteps)){
      popSizeCom <- rbind(popSizeCom,ricker(popSizeCom[i,],params$r,params$carryingCapacity,eps_com[i,]))
      popSizeRes <- rbind(popSizeRes,ricker(popSizeRes[i,],params$r,params$carryingCapacity,eps_res[i,]))
      popSizeNewRes <- rbind(popSizeNewRes,ricker(popSizeNewRes[i,],params$r,params$carryingCapacity,eps_new_res[i,]))
  }
  
  totalPopSizeCom <- rowSums(popSizeCom)
  totalPopSizeRes <- rowSums(popSizeRes)
  totalPopSizeNewRes <- rowSums(popSizeNewRes)
  
  if(extinct(totalPopSizeCom,tLow,tHigh)){numExtinctCom<-numExtinctCom+1}
  if(extinct(totalPopSizeRes,tLow,tHigh)){numExtinctRes<-numExtinctRes+1}
  if(extinct(totalPopSizeNewRes,tLow,tHigh)){numExtinctNewRes<-numExtinctNewRes+1}
   
  if(j == 1){
    plot.ts(totalPopSizeCom)
    plot.ts(totalPopSizeRes)
    plot.ts(totalPopSizeNewRes)

    hist(totalPopSizeCom,200)
    hist(totalPopSizeRes,200)
    hist(totalPopSizeNewRes,200)
  }
}
```


